Tree regeneration in models of forest dynamics: a key priority for further research

### 2023-04-28 manuscript/manuscriptRW.Rmd
changes from Harald

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbances impacts and changing societal
demands on forests ecosystems services, it is becoming ever more
important to understand how forest structure, composition and function
will change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics were developed over the past decades considering the
impacts of climate on forest dynamics [@bugmann2022]. From these
studies, it is evident that we have a reasonably good understanding how
to model tree growth [@vanclay1997; @bugmann1996], and substantial
efforts have been dedicated to improving the representation of tree
mortality [@bugmann2019; @cailleret2017]. In contrast, tree regeneration
is a much less studied process that is often represented rather coarsely
in these models [@walck2011; @leishman1992; @price2001], which is a
clear research gap in the context of climate-induced forest disturbances
and forest resilience.

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994].

Overall, models are needed to (1) synthesize existing empirical data and
explore their relationships, (2) assess future tree regeneration e.g.,
in the context of global change scenarios, and (3) identify the most
important processes that are shaping ecological patterns. Given the
current strategies that are used in models of forest dynamics to
represent tree regeneration, model behavior often is prone to problems,
such as very high levels of tree regeneration that necessitate excess
mortality at early stages of tree life to simulate correct stand
structure and composition. Also, inadequately high species diversity in
tree regeneration may be simulated, which is characteristic of
"classical" forest gap models [@gutiérrez2016], at least as long as the
simulations set-up comprise a multitude of species. Some models use
calibration against local canopy-level data to define appropriate
regeneration levels, which is likely to strongly hamper the general
applicability of these models e.g. under scenarios of climate change.
Furthermore, correctly capturing the species composition in tree
regeneration as a function of the presence of seed trees in the canopy
is often a particular challenge, potentially leading to unrealistic
successional drift in the model, which must be corrected by factors that
are hard to parameterize [e.g., @lischke2006]. A related issue is the
excessive reduction of species diversity due to positive feedback
effects such that eventually just single-species stands remain. This is
sometimes corrected by the incorporation of a low level of seed influx
of all species at all times [@schumacher2006] or by restricting the
number of seeds per species in the seedbank [@lischke2006]. However,
simulated species composition is usually exceedingly sensitive to
assumptions about seed availability, whereas the parameters of such
functions are poorly constrained by field data. Lastly, there is often a
problem with insufficient observational constraints on parameter values
for models that start from very small tree sizes (e.g., 10 cm of height
or even seeds) and track tree development in a process-oriented manner
by considering a multitude of ecological influences, rather than
emphasizing tree regeneration into a larger size class [e.g.,
@wehrli2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. Also in empirical
ecology, tree regeneration has not received a lot of attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data is available that covers all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Thus, the available data are fragmented,
which constitutes a major problem for model building [@clark1999]. On
the one hand, monitoring on permanent sample plots (such as National
Forest Inventories) often measures tree regeneration [@zell2019],
however with a design that captures the rate after the stand initiation
phase [@hallsby2015]. Data from permanent plots, although available
across regions, are highly heterogeneous, rendering their use difficult
in a modeling context [@käber2021]. On the other hand, targeted studies
to measure tree regeneration on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2022]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
in tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated in
most models such as deer browsing --- we do not aim at a detailed
statistical evaluation of each model. Instead, we aim at evaluating the
general regeneration patterns and magnitudes simulated by the models,
and benchmark the simulated regeneration niche of multiple species
against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

@@

Fifteen models of forest dynamics (including two models featuring model
variants) were used to assess simulated tree regeneration. They simulate
forest dynamics at the stand, landscape or global scale (Table
\@ref(tab:tableModels)). The approaches used for model construction and
their origin differ strongly, with most of the models featuring a
largely "process-based" approach, whereas two models are based on
formulations derived from the statistical analysis of inventory data
(SIBYLA and xComp). Some of the models largely rely on the original
approach underlying forest gap models (e.g., ForClim 1 and LandClim),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree regeneration processes.

The regeneration modules has been categorized as 'regeneration' or
'recruitment' models [@vanclay1994; @konig_tree_2022]. *Regeneration
models* include processes such as flowering and pollination, seed
production, seed dispersal, germination, and seedling growth, which
ultimately lead to the simulated number of established trees.
*Recruitment models*, in contrast, introduce a number of new trees with
certain characteristics such as biomass or diameter, without explicitly
considering earlier development processes. We can further distinguish
models that feature a feedback in the regeneration, i.e., where the
level and species composition of regeneration is influenced by the
existing adult tree community via the production of seeds, seedlings or
saplings, from models that do not contain such a feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes which are not treated explicitly. This later approach reduces
model complexity but comes at the cost of blurring process
representation.

In the models used in this study, the overall complexity in the
regeneration modules varies considerably. Following @bugmann2022, we can
classify seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS,
ForClim and TreeMig) as having rather high complexity in their
regeneration modules (mean regeneration formulation complexity across
all processes \>0, Table \@ref(tab:tableModels)). Two models, ForClim
and ForCEEPS, were used here with two alternative variants of
regeneration. ForClim variant 1 [@bugmann1996] is based on a recruitment
module that adheres closely to the concept introduced by @botkin1972,
whereas ForClim variant 11 is adopting a different approach where
individual species properties and their relation to the environment are
important [@huber2020]. These two model variants allow us to evaluate a
more process-based and complex module (variant 11) against a simple
module (variant 1), while the rest of the model structure is identical.
Similarly, the two ForCEEPS variants allow us to isolate the importance
of the canopy feedback (i.e., simulated actual composition and relative
abundance of species in the plot) via seed trees for the quantity and
quality (e.g., diversity and composition) of simulated regeneration, as
one variant includes this feedback whereas the other does not.

@@

Regeneration data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel and unprecedented network of sites in forest reserves that
represent the range of environmental gradients in temperature and
precipitation in Central Europe as compiled in the framework of the
EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
have been unmanaged for at least 10 years prior to their designation as
reserves, and most of them provide time series of natural forest
dynamics over multiple decades and up to ca. 80 years. The census
periods range from 3 to 37 years, with an average of 14 years. The data
provide information at the tree level, thus allowing for sequential
comparison such as individual-tree regeneration and death between the
measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation. The selection of sites is representative of the
environmental variation of the data. This was achieved by applying
k-means clustering to define 200 clusters of plots from the original set
of 869 plots along the environmental dimensions of temperature, climatic
water balance, soil quality and exposition (i.e., slope and aspect).
Each of these sites featured at least two consecutive measurements.
egeneration thresholds for the sites differed between diameters of 0 and
10 cm. For our study, we defined two datasets where one included 165
sites with a diameter threshold of 7 cm or lower, and the other included
another 35 sites with diameter thresholds between 7 and 10 cm.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2022 to provide regeneration
rates per ha and per decade. The observations used in this study
featured 30,900 established trees. Regeneration rates per site, sample
and decade ranged from zero to 1246 trees per ha, with a mean of 56
trees per ha. Adult species composition was also available at the plots.
For more details on this unique dataset and the detailed regeneration
information, cf. @käber2022.

@@

The overarching goal of the experiments was to assess the tree
regeneration as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define the tree
regeneration as the passing of a breast height diameter threshold of 7
or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
group was provided with a detailed protocol [@díaz-yáñez2022] with
instructions for how to perform the simulations, providing input
variables on climate and soil conditions, and the list of expected
output variables. Neither were further site information (except for the
data specified below) nor any data on tree regeneration or forest stand
features provided prior to the simulation. That is, the models were run
in "blind flight" mode.

The input variables were collected from different data sources and
aggregated to be adapted to the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data was different in each model, based on the
approach that best suited the model (Table \@ref(tab:tableModels)). Soil
quality data were provided as continuous values between 1 and 5
[Soilgrids dataset, Hengl et al. -@hengl2017]. The protocol also
provided the elevation, slope and aspect for each of the 200 sites
[@asterscienceteam2019], but no other spatial information such as
coordinates, with the exception of iLand and aDGVM2, which required
blurred coordinates to derive highly detailed soil data.

@@

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree regeneration under comparable
and near equilibrium conditions. This entails the assumption that the
observations from the forest reserves (1) reflect no traces of forest
management and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of the regeneration process (i.e., the number of ingrowth
trees per area and per unit of time) were of interest.

The simulations were run for mixed species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
parameterised. XComp simulations did not consider *Fraxinus excelsior*
due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
pseudoplatanus* and *Alnus glutinosa* are not parameterized and
therefore these species could not be included in the simulation.
Finally, the model aDGVM2 does not simulate individual species. Rather,
community assembly processes and trait filtering generate plant
communities that are adjusted to the biotic and abiotic conditions, and
the simulated plants can be classified into ecological strategies based
on their trait values in a post-processing step.

Each of the models reported the regeneration number by sampling 200
times in a 10-year interval for each species and per ha for each of the
200 sites. Multiple samples per site were used to better understand the
simulated variation within each site. This was done using different
strategies, depending on the model, including (1) sampling simulated
data from the same 1 ha plot in the equilibrium over time, (2) sampling
several 1 ha plots from the simulated forest at one specific point in
time (in the equilibrium) or (3) a combination of (1) and (2) . This
resulted in 880,000 observations per model that simulated the 11 species
included in the protocol (200 sites, 200 samples per site, 11 species
and two diameter thresholds). For the models that simulated additional
species, their regeneration rates were aggregated as "others".

@@

The simulation results were analyzed regarding (1) regeneration levels
(i.e., ingrowth number per unit time and space), (2) regeneration
species diversity, (3) regeneration mortality, (4) model performance and
model traits, and (5) ingrowth gradients along regeneration niches. We
evaluated the species diversity across models and in relation to the
observed data by calculating the Shannon index based on the relative
proportion of the species in terms of basal area. The Shannon index was
calculated for the regeneration ($H_{R_{n}}$) (Equation
\ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$) (Equation
\ref{eqpiShannon}). The higher the value of the index, the higher the
species diversity at a particular site and sample.

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present and that have a basal
area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
the species $i$ in sample $n$ calculated as the regeneration basal area
($r_{BA_{n,i}}$) for that species $i$ relative to the total recruited
basal area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of the species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to the total basal area
($TotalR_{BA_{n}}$) of the sample $n$. The species diversity was not
assessed for the model aDGVM2 because it does not simulate individual
species.

Mortality in tree regeneration was assessed based on the ratio of
regeneration between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
to estimate whether the ratio of regeneration between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate
between these two diameters. The Reineke self-thinning rule is usually
calculated from even-aged, single species stands and assumes a fixed
relationship between the number of stems and the quadratic mean diameter
in fully stocked pure stands. The value used in our comparisons was 1.77
(i.e., we expect stem numbers at 7 cm to be 77% higher than at 10 cm),
calculated using Equation \ref{eqReineke}.

@@

Total regeneration and the regeneration niches of the individual species
were evaluated across the environmental gradients of light availability,
temperature, and soil moisture. Stand basal area (the basal area of all
the trees in each sample including tree regeneration) was used as a
proxy of light availability at the forest floor, the annual degree-day
sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
and the climatic water balance as a proxy for soil moisture
[@speich2019]. Regeneration values were calculated as the mean across
the 200 samples per site. The observed data were modeled using a
Generalized Additive Model [@wood2011] with a negative binomial
distribution and restricted maximum likelihood to better understand the
relationship between the environmental gradients and the levels of
observed regeneration, relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic
gradients, we focused on the five main tree species or genera: *Abies
alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
*Quercus* spp. For these species, we calculated the share in the
regeneration basal area per site as the mean across the available
samples per site:

>>>>

Forests provide a wide range of products and services of vital
importance to humankind [@foresteurope2020]. Under the pressure of
climate change, increasing disturbance impacts and changing societal
demands on forest ecosystem services, it is becoming ever more important
to understand how forest structure, composition and function will
change, and to evaluate forest capacity to adapt to or be resilient
after disturbance [@lindner2010; @seidl2022]. A wide range of models of
forest dynamics were developed over the past decades considering the
impacts of climate [@bugmann2022]. From these studies, it is evident
that we have a reasonably good understanding how to model tree growth
[@vanclay1997; @bugmann1996], and substantial efforts have been
dedicated to improving the representation of tree mortality
[@bugmann2019; @cailleret2017]. In contrast, tree regeneration is a much
less studied process that is often represented rather coarsely in models
[@walck2011; @leishman1992; @price2001]. Which is a clear research gap
in the context of climate-induced forest disturbances and forest
resilience.

Tree regeneration arises from multiple processes such as pollination,
fruit maturation, seed production, dispersal, germination, juvenile
growth and survival [@price2001; @vacchiano2018]. All these processes
are difficult to assess, and some of them are scarcely understood and
thus appear highly stochastic [@bogdziewicz2021]. Identifying the
appropriate level of complexity for the mathematical formulation of the
key factors that are leading to successful tree regeneration is
challenging. Currently, tree regeneration processes in dynamic forest
models are handled in a multitude of ways [@konig_tree_2022;
@bugmann2022]: from 1) entirely ignoring it [as done in classical forest
growth models, e.g., Pretzsch et al. -@pretzsch2002], across 2) the use
of a few simple environmental filters, as done in most forest "gap"
models [@shuman2011] and Dynamic Global Vegetation Models [e.g.,
@hickler2012; @smith2001], to 3) complex approaches that incorporate
local feedback from the canopy, multiple ecological processes and often
also short time steps [e.g., @seidl2012; @wehrli2006], or 4) field-based
statistical parameterizations, which however are not easy to extrapolate
in space and time [e.g., @ribbens1994]. Overall, models are needed to
(1) synthesize existing empirical data and explore their relationships,
(2) assess future tree regeneration e.g. in the context of global change
scenarios, and (3) identify the most important processes that are
shaping ecological patterns. Given the current strategies that are used
in models of forest dynamics to represent tree regeneration, model
behavior often is prone to problems, such as very high levels of tree
regeneration that necessitate excess mortality at early stages of tree
life to simulate correct stand structure and composition. Also,
inadequately high species diversity in tree regeneration may be
simulated, which is characteristic of "classical" forest gap models
[@gutiérrez2016], at least as long as the simulation set-up comprises a
multitude of species. Some models use calibration against local
canopy-level data to constrain simulated regeneration levels, which is
likely to hamper the general applicability of these models e.g. under
scenarios of climate change. Furthermore, correctly capturing the
species composition in tree regeneration as a function of the presence
of seed trees in the canopy is often a particular challenge, potentially
leading to unrealistic successional drift in the model, which must be
corrected by factors that are hard to parameterize [e.g., @lischke2006].
A related issue is the excessive reduction of species diversity due to
positive feedback effects, such that eventually just single-species
stands remain. This is sometimes corrected by the incorporation of a low
level of seed influx of all species at all times [@schumacher2006] or by
restricting the number of seeds per species in the seedbank
[@lischke2006]. However, simulated species composition is usually
exceedingly sensitive to assumptions about seed availability, whereas
the parameters of such functions are poorly constrained by field data.
Lastly, there is often a problem with insufficient observational
constraints on parameter values for models that start from very small
tree sizes (e.g., 10 cm of height or even seeds) and track tree
development in a process-oriented manner by considering a multitude of
ecological influences, rather than emphasizing tree regeneration into a
larger size class [e.g., @wehrli2006].

Thus, a focus on the modeling of tree regeneration processes is sorely
needed and overdue [@price2001; @walck2011] if we are to make reliable
projections of future forest dynamics, i.e., when the models need to be
operated in extrapolation mode, as well as from a fundamental ecological
point of view for increasing systems understanding. Also in empirical
ecology, tree regeneration has not received a lot of attention
[@hanbury-brown2022], although many sub-processes have been studied in
detail [@miina2006]. Yet, little data are available that covers all the
processes within one species along environmental gradients, let alone
for a vast suite of species. Thus, the available data are fragmented,
which constitutes a major problem for model building [@clark1999]. On
the one hand, monitoring on permanent plots (such as National Forest
Inventories) often measures tree regeneration [@zell2019], however with
a design that captures the rate after the stand initiation phase
[@hallsby2015]. Data from permanent plots, although available across
regions, are highly heterogeneous, rendering their use difficult in a
modeling context [@käber2021]. On the other hand, targeted studies to
measure tree regeneration on experimental sites are often limited in
spatial extent, species studied, or the sub-set of processes that are
targeted [@berdanier2016; @collet2006].

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree regeneration [@käber2022]. By tree regeneration, we refer to the
passing of trees across a specific diameter threshold ("ingrowth"). We
included models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Due to the large variability in
tree regeneration patterns in nature and the large number of factors
driving this process --- including some that are not incorporated
explicitly in most models, such as deer browsing --- we do not aim for a
detailed statistical evaluation of each model. Instead, we aim to
evaluate the general regeneration patterns and magnitudes simulated by
the models, and benchmark the simulated regeneration niche of multiple
species against empirical data along a wide environmental gradient of
temperature, moisture, and light availability [@grubb1977].

@@

Fifteen models of forest dynamics (including two models featuring model
variants) were used to simulate forest dynamics at the stand, landscape
or global scale (Table \@ref(tab:tableModels)). The approaches used for
model construction and their origin differ strongly, with most of the
models featuring a largely "process-based" approach, whereas two models
are based on formulations derived from the statistical analysis of
inventory data (SIBYLA and xComp). Some of the models largely rely on
the original approach underlying forest gap models (e.g., ForClim 1),
whereas some are based on plant ecophysiological processes (e.g.,
FORMIND and iLand). The different approaches underlying the models have
strongly influenced the formulation of tree regeneration processes.

The regeneration models can be differentiated into 'regeneration' and
'recruitment' approaches [@vanclay1994; @konig_tree_2022]. *Regeneration
models* include processes such as flowering and pollination, seed
production, seed dispersal, germination, and seedling growth, which
ultimately lead to the simulated number of established trees.
*Recruitment models*, in contrast, introduce a number of new trees with
certain characteristics such as biomass or diameter, without explicitly
considering earlier development processes. We can further distinguish
models that feature a feedback in the regeneration, i.e., where the
level and species composition of regeneration is influenced by the
existing adult tree community via the production of seeds, seedlings or
saplings, from models that do not contain such a feedback.

The starting point for tree regeneration in the models differs as well,
ranging from seedbank, seed or seedling to sapling (i.e., trees much
larger than 10 cm height, often ca. 2-3 m tall). Models that start from
seed need to include a larger number of ecological processes such as
germination and survival of young seedlings, whereas models that start
from saplings have to aggregate via parameterizations several ecological
processes that are not treated explicitly. This latter approach reduces
model complexity but comes at the cost of blurring process
representation.

In the models used here, the overall complexity in the regeneration
modules varies considerably. Following @bugmann2022, we can classify
seven models (iLand, PICUS, LandClim, ForCEEPS, LPJ-GUESS, ForClim and
TreeMig) as having rather high complexity in their regeneration modules
(mean regeneration formulation complexity across all processes \>0,
Table \@ref(tab:tableModels)). Two models, ForClim and ForCEEPS, were
used here with two alternative variants of regeneration. ForClim variant
1 [@bugmann1996] is based on a recruitment module that adheres closely
to the concept introduced by @botkin1972, whereas ForClim variant 11 is
adopting a different approach where individual species properties and
their relation to the environment are important [@huber2020]. These two
model variants allow us to evaluate a more process-based and complex
module (variant 11) against a simple module (variant 1), while the rest
of the model structure is identical. Similarly, the two ForCEEPS
variants allow us to isolate the importance of the canopy feedback
(i.e., simulated actual composition and relative abundance of species in
the plot) via seed trees for the quantity and quality (e.g., diversity
and composition) of simulated regeneration, as one variant includes this
feedback whereas the other does not.

@@

Regeneration data covering a wide range of environmental conditions are
hard to obtain, and this is one of the reasons why most models of forest
dynamics have never been confronted with a dataset covering such
gradients over a large number of sites, to evaluate how well
regeneration is captured. The observations used here are derived from a
novel and unprecedented network of sites in forest reserves that
represent the range of environmental gradients in temperature and
precipitation in Central Europe as compiled in the framework of the
EuFoRIa network [@euforia2019] (Figure \@ref(fig:map)). These forests
have been unmanaged for at least 10 years prior to their designation as
reserves, and most of them provide time series of natural forest
dynamics over multiple decades and up to ca. 80 years. The census
periods range from 3 to 37 years, with an average of 14 years. The data
provide information at the tree level, thus allowing for the sequential
comparison of processes such as individual-tree regeneration and death
between the measurements.

We selected 200 sites from this network as the benchmarking dataset for
the simulation to be representative of the environmental variation
contained in the data. This was achieved by applying k-means clustering
to define 200 clusters of plots from the original set of 869 plots along
the environmental dimensions of temperature, climatic water balance,
soil quality and exposition (i.e., slope and aspect). Each of these
sites featured at least two consecutive measurements. Regeneration
thresholds for these sites differed between diameters of 0 and 10 cm.
For our study, we defined two datasets where one included 165 sites with
a diameter threshold of 7 cm or lower, and the other included another 35
sites with diameter thresholds between 7 and 10 cm.

Plot size ranged from 0.02 ha to 5.52 ha, and the data were further
processed and aggregated following @käber2022 to provide regeneration
rates per ha and per decade. The observations used in this study
featured 30,900 newly established trees. Regeneration rates per site,
sample and decade ranged from zero to 1246 trees, with a mean of 56
trees. Adult species composition was also available for each plot. For
more details on this unique dataset and the detailed regeneration
information, cf. @käber2022.

@@

The overarching goal of the experiments was to assess the tree
regeneration as it arises from empirical data against its representation
in a wide range of models of forest dynamics. We define tree
regeneration as the passing of a breast height diameter threshold of 7
or 10 cm, respectively (synonym: ingrowth). To this end, each modeling
group was provided with a detailed protocol [@díaz-yáñez2022] with
instructions how to perform the simulations, providing input variables
on climate and soil conditions, and the list of expected output
variables. Neither were further site information (except for the data
specified below) nor any data on tree regeneration or forest stand
features provided prior to the simulation. That is, the models were run
in "blind flight" mode.

The input variables were collected from different data sources and
aggregated to be adapted to the needs of the different models. Time
series of climatic variables from 1981 to 2018 were provided in hourly
[Era5-land data, Muñoz Sabater -@muñozsabater2019], daily and monthly
resolution [CHELSA data, Karger et al. -@karger2021]. Some variables
required for some models, such as relative humidity or vapor pressure
deficit, were calculated from these variables. The final instructions
for using the climate data were slightly different in each model, based
on the approach that best suited the model (Table
\@ref(tab:tableModels)). Soil quality data were provided as continuous
values between 1 and 5 [Soilgrids dataset, Hengl et al. -@hengl2017].
The protocol also provided the elevation, slope and aspect for each of
the 200 sites [@asterscienceteam2019], but no other spatial information
such as coordinates, with the exception of iLand and aDGVM2, which
required blurred coordinates to derive highly detailed soil data.

@@

The simulations were run in the absence of management to a simulated
equilibrium ("Potential Natural Vegetation") with the current climate,
as the ultimate goal was to evaluate tree regeneration under comparable
and near-equilibrium conditions. This entails the assumption that (1)
the observations from the forest reserves reflect no traces of forest
management, and (2) there is an equilibrium between forest dynamics and
climate. While the former might be starting to be visible in many of the
EuFoRIA reserves, the latter may be more debatable. However, in the
absence of detailed data on the history of each plot in the EuFoRIa
network, some broad assumptions had to be made. Both the width of the
regeneration niche (i.e., in environmental space) as well as the
intensity of the regeneration process (i.e., the number of ingrowth
trees per area and per unit of time) were of interest.

The simulations were run for mixed-species forests (not multiple
single-species simulations) using mixtures of eleven species or genera
for which regeneration data of sufficient quality were available from
EuFoRIa: *Fagus sylvatica L., Picea abies L., Abies alba Mill., Carpinus
betulus L., Tilia cordata Mill., Acer pseudoplatanus L., Betula spp. L.,
Fraxinus excelsior L., Quercus spp. L., Alnus glutinosa L., Pinus
sylvestris L.* The same set of species was used at all 200 sites. Two
models included their standard set of species for the simulations, which
is much larger (i.e., ForClim 1, ForClim 11, TreeMig). In three models,
fewer than these eleven species were simulated (4C, xComp, LPJ-GUESS)
(Table \@ref(tab:tableModels)). In 4C only *Fagus sylvatica*, *Picea
abies*, *Betula* spp., *Quercus* spp. and *Pinus sylvestris* are
parameterised. xComp simulations did not consider *Fraxinus excelsior*
due to a depreciated species parameterisation. In LPJ-GUESS, *Acer
pseudoplatanus* and *Alnus glutinosa* are not parameterized and
therefore these species could not be included in the simulation.
Finally, the model aDGVM2 does not simulate individual species. Rather,
community assembly processes and trait filtering generate plant
communities that are adjusted to the biotic and abiotic conditions, and
the simulated plants can be classified into ecological strategies based
on their trait values in a post-processing step.

Each of the models reported the regeneration number by sampling 200
times in a 10-year interval for each species and per ha for each of the
200 sites. Multiple samples per site were used to better understand the
simulated variation within each site. This was done using different
strategies, depending on the model, including (1) sampling simulated
data from the same 1 ha plot in the equilibrium over time, (2) sampling
several 1 ha plots from the simulated forest at one specific point in
time (in the equilibrium) or (3) a combination of (1) and (2). This
resulted in 880,000 observations per model that simulated the 11 species
included in the protocol (200 sites, 200 samples per site, 11 species
and two diameter thresholds). For the models that simulated additional
species, their regeneration rates were aggregated as "others".

@@

The simulation results were analyzed regarding (1) regeneration levels
(i.e., ingrowth number per unit time and space), (2) regeneration
species diversity, (3) regeneration mortality, (4) the relationship
between model performance and model traits, and (5) ingrowth gradients
along the regeneration niches. We evaluated species diversity across the
models and in relation to the observed data by calculating the Shannon
index $H$ based on the relative proportion of the species in terms of
basal area. It was calculated for the regeneration ($H_{R_{n}}$)
(Equation \ref{eqpiShannonR}) and at the stand level ($H_{S_{n,s}}$)
(Equation \ref{eqpiShannon}). The higher the value of the index, the
higher the species diversity at a particular site and sample.

```{=tex}
\begin{equation}
H_{R_{n}} = - \sum_{i = 1}^{s}{p_{R_{n,i}}\ln(p_{R_{n,i}})}\label{eqpiShannonR}
\end{equation}
```
```{=tex}
\begin{equation}
p_{R_{n,i}} = \frac{r_{BA_{n,i}}}{TotalR_{BA_{n}}} \label{eqpiShannonR2}
\end{equation}
```
```{=tex}
\begin{equation}
H_{S_{n}} = - \sum_{i = 1}^{s}{p_{S_{n,i}}\ln(p_{S_{n,i}})}\label{eqpiShannon}
\end{equation}
```
```{=tex}
\begin{equation}
p_{S_{n,i}} = \frac{BA_{n,i}}{Total_{BA_{n}}} \label{eqpiShannon2}
\end{equation}
```
where $s$ is the total number of species present and that have a basal
area larger than zero in sample $n$; $p_{R_{n,i}}$ is the proportion of
species $i$ in sample $n$ calculated as the regeneration basal area
($r_{BA_{n,i}}$) for that species $i$ relative to total recruited basal
area ($TotalR_{BA_{n}}$) of the sample $n$; $p_{S_{n,i}}$ is the
proportion of species $i$ calculated as the basal area of all trees
($BA_{n,i}$) of that species $i$ relative to total basal area
($TotalR_{BA_{n}}$) of the sample $n$. Species diversity was not
assessed for aDGVM2 as this model does not simulate individual species.

Mortality in tree regeneration was assessed based on the ratio of
regeneration between the 7 and 10 cm diameter thresholds. We used the
Reineke self-thinning rule [@reineke1933; @pretzsch2003] as a reference
to estimate whether the ratio of regeneration between the 7 and 10 cm
diameter thresholds were above or below the expected theoretical rate.
The Reineke self-thinning rule is usually calculated for even-aged,
single species stands and assumes a fixed relationship between the
number of stems and the quadratic mean diameter in fully stocked pure
stands. The value used in our comparisons was 1.77 (i.e., we expect stem
numbers at 7 cm to be 77% higher than at 10 cm), calculated using
Equation \ref{eqReineke}.

@@

Total regeneration and the regeneration niches of the individual species
were evaluated across the environmental gradients of light availability,
temperature, and soil moisture. Stand basal area (the basal area of all
the trees in each sample including tree regeneration) was used as a
proxy for light availability at the forest floor, the annual degree-day
sum [@allen1976; @fischlin1995] as a proxy for growing season warmth,
and the climatic water balance as a proxy for soil moisture
[@speich2019]. Regeneration values were calculated as the mean across
the 200 samples per site. The observed data were modeled using a
Generalized Additive Model [@wood2011] with a negative binomial
distribution and restricted maximum likelihood to better understand the
relationship between the environmental gradients and the levels of
observed regeneration, relative to the simulation results per model.

In order to analyze the regeneration niches across the climatic
gradients, we focused on five common tree species or genera: *Abies
alba*, *Fagus sylvatica*, *Picea abies*, *Pinus sylvestris*, and
*Quercus* spp. For these species, we calculated the share in the
regeneration basal area per site as the mean across the available
samples per site:
